---
title: "Case Study 1; Task 1: Fibonacci numbers"
author: "Jonas Tatzberger"
date: "2024-04-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# a) Defining functions with for & while loops
At first the function calculating the ratio using a for loop:
```{r, echo=TRUE}
r_for <- function(i) {
  f_i <- 0
  f_ip1 <- 1
  result <- c()
  for (x in 1:i) {
    f_ip1_new <- f_i+f_ip1
    f_i <- f_ip1
    f_ip1 <- f_ip1_new
    result <- append(result, f_ip1/f_i)
  }
  return (result)
}
```
Running the function using a test output:
```{r, echo=TRUE}
test_for <- r_for(10)
print(test_for)
```
Next the function using a while-loop:
```{r, echo=TRUE}
r_while <- function(i) {
  f_i <- 0
  f_ip1 <- 1
  result <- c()
  x=0
  while(x < i) {
    f_ip1_new <- f_i+f_ip1
    f_i <- f_ip1
    f_ip1 <- f_ip1_new
    result <- append(result, f_ip1/f_i)
    x = x+1
  }
  return (result)
}
```
Running the function using a test output:
```{r, echo=TRUE}
test_while <- r_while(10)
print(test_while)
```
As we can see, both functions produce the same outputs, which seem to nicely converge towards the golden ratio.

# b) Benchmarking the function to see which is more efficient
To test the functions, the microbenchmark package is used. When applying the function, the standard values are used (the number of iterations for the expression is set to 100 here).
This results in a statistic comprised of n=100 tests.
```{r, echo=TRUE}
library(microbenchmark)
```
## Testing for n=200: 
```{r, echo=TRUE}
microbenchmark(r_for(200))
microbenchmark(r_while(200))
```
Here we can see that the mean time that the "r_for" function took (397.028us) is lower than the "r_while" function (423.145us). For n=200 the for loop is faster.

## Testing for n=2000:
```{r, echo=TRUE}
microbenchmark(r_for(2000))
microbenchmark(r_while(2000))
```
Again we can see that the mean time that the "r_for" function took (18.295ms) is lower than the "r_while" function (20.847ms). For n=2000 also the for-loop is faster.

# c) Plotting for n=100:
In order to visualize the function, the package "ggplot2" is used. To work with the data it is transformed into a data-frame and an iteration column is added.
```{r, echo=TRUE}
library(ggplot2)
value = r_for(100)
r_100 <- data.frame(value)
r_100$iteration <- 1:nrow(r_100)
ggplot(r_100, aes(x=iteration, y=value)) + geom_point() + geom_line() + ggtitle("Visualization of the approach to the golden ratio with n=100 iterations")
```
From the plot, one can deduce that the function converges at around the 6th iteration, from this point on no significant changes are visible.


