```{r import_exercise3, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}

import_library <- function(library_name) {
  if (!requireNamespace(library_name, quietly = TRUE)) {
    install.packages(library_name)
  }
  library(library_name, character.only = TRUE)
}

# --------------------------

import_library("ggplot2")
import_library("readr")
import_library("dplyr")
import_library("tidyr")
import_library("tidyverse")
import_library("knitr")
import_library("plotly")

```

```{r path_determination_exercise3, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

```{r prepare_data_e3, include=FALSE}

data_tibble <- read_delim(file = paste0(working_directory_path, "/data/final_data.csv"),
                          show_col_types = FALSE)

data_tibble <- data_tibble %>% 
  filter(Classification_2020 != ".")

data_tibble <- data_tibble %>% 
  mutate(Classification_2020 = factor(Classification_2020, 
                                      levels = c("L", "LM", "UM", "H"), 
                                      labels = c("Low", "Lower-middle", "Upper-middle", "High"), 
                                      ordered = TRUE)
  )

data_tibble <- data_tibble %>%
  mutate(
    Continent = factor(Continent),
    Region = factor(Region)
  )

data_tibble <- data_tibble %>%
 mutate(
    Net_Migration_Rate = na_if(Net_Migration_Rate, ".") %>% as.double(),
    Median_Age = na_if(Median_Age, ".") %>% as.double(),
    Youth_Unemployment_Rate = na_if(Youth_Unemployment_Rate, ".") %>% as.double()
  )

head(data_tibble)

```

## k. Median net migration rate per subcontinent.

Go online and find a data set which contains the 2020 population for the 
countries of the world together with ISO codes.

 - Download this data and merge it to the dataset you are working on in this 
   case study using a left join. (A possible source: [World Bank ](https://data.worldbank.org/indicator/SP.POP.TOTL?end=2020&start=2020))
 - Inspect the data and check whether the join worked well.

**Answer**

```{r k1, dev=c('png')}

```

## l. Scatterplot of median age and net migration rate in Europe

Make a scatterplot of median age and net migration rate for the countries of Europe. 
Scale the size of the points according to each country’s population.

 - For better visibility, use a transparency of alpha=0.7.
 - Remove the legend.
 - Comment on the plot.

**Answer**

## m. Interactive plot

On the merged data set from Task k., using function ggplotly from package plotly
re-create the scatterplot in Task l., but this time for all countries. 
Color the points according to their continent.

When hovering over the points the name of the country, the values for median age, 
net migration rate, and population should be shown. 
(Hint: use the aesthetic text = Country. 
In ggplotly use the argument tooltip = c("text", "x", "y", "size")).

**Answer**

## n. Parallel coordinate plot

In parallel coordinate plots each observation or data point is depicted as a line 
traversing a series of parallel axes, corresponding to a specific variable or dimension. 
It is often used for identifying clusters in the data.

One can create such a plot using the GGally R package. You should create such a 
plot where you look at the three main variables in the data set: median age, 
youth unemployment rate and net migration rate. 
Color the lines based on the income status. Briefly comment.

**Answer**

## o. World map visualisation

Using the package rworldmap, create a world map of the median age per country. 
Use the [vignette](https://cran.r-project.org/web/packages/rworldmap/vignettes/rworldmap.pdf) to find how to do this in R.

**Answer**

