```{r import_exercise2, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}

import_library <- function(library_name) {
  if (!requireNamespace(library_name, quietly = TRUE)) {
    install.packages(library_name)
  }
  library(library_name, character.only = TRUE)
}

# --------------------------

import_library("ggplot2")
import_library("readr")
import_library("dplyr")
import_library("tidyr")
import_library("tidyverse")
import_library("knitr")
import_library("forcats")

```

```{r path_determination_exercise2, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

```{r prepare_data_e2, include=FALSE}

data_tibble <- read_delim(file = paste0(working_directory_path, "/data/final_data.csv"),
                          show_col_types = FALSE)

data_tibble <- data_tibble %>% 
  filter(Classification_2020 != ".")

data_tibble <- data_tibble %>% 
  mutate(Classification_2020 = factor(Classification_2020, 
                                      levels = c("L", "LM", "UM", "H"), 
                                      labels = c("Low", "Lower-middle", "Upper-middle", "High"), 
                                      ordered = TRUE)
  )

data_tibble <- data_tibble %>%
  mutate(
    Continent = factor(Continent),
    Region = factor(Region)
  )

data_tibble <- data_tibble %>%
 mutate(
    Net_Migration_Rate = na_if(Net_Migration_Rate, ".") %>% as.double(),
    Median_Age = na_if(Median_Age, ".") %>% as.double(),
    Youth_Unemployment_Rate = na_if(Youth_Unemployment_Rate, ".") %>% as.double()
  )

head(data_tibble)

```

## f. Median net migration rate per subcontinent.

The plot in task e. shows the distribution of the net migration rate for each subcontinent. 
Here you will work on visualizing only one summary statistic, namely the median.
For each subcontinent, calculate the median net migration rate. Then create a plot which contains the
sub-regions on the y-axis and the median net migration rate on the x-axis.

 - As geoms use points.
 - Color the points by continent – use a colorblind friendly palette 
   (see e.g., [here](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)).
 - Rename the axes.
 - Using fct_reorder from the forcats package, arrange the levels of subcontinent 
   such that in the plot the lowest (bottom) subcontinent contains the lowest 
   median net migration rate and the upper most region contains the highest median 
   net migration rate.
 - Comment on the plot. E.g., what are the regions with the most influx? 
   What are the regions with the most outflux?

**Answer**

```{r f1, dev=c('png')}
data_tibble <- data_tibble %>% filter(!is.na(Youth_Unemployment_Rate))
median_migration <- data_tibble %>%
  group_by(Region, Continent) %>%
  summarize(median_net_migration_rate = median(Net_Migration_Rate)) %>%
  ungroup()

median_migration <- median_migration %>%
  mutate(Region = fct_reorder(Region, median_net_migration_rate, .na_rm = TRUE))

ggplot(median_migration, aes(x = median_net_migration_rate, y = Region, color = Continent)) +
  geom_point(size = 4, na.rm = TRUE) +
  scale_color_brewer(palette = "Set2") +
  labs(
    x = "Median Net Migration Rate",
    y = "Subcontinent",
    title = "Median Net Migration Rate by Subcontinent",
    color = "Continent"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))
```

This plot shows that eastern regions like Central Asia, Polynesia and Micronesia show the most outflux, with the latter islands even more seperated from the pack. Most African regions show a balanced net migration rate. North America, large parts of Europe and Australia are the continents with the largest influx, with Australia heavily dominating this statistic. Given the geographic proximity of Australia and Micronesia and Polynesia, it seems likely that a large amount of the outflux of the latter migrate to Australia and thus explain the median net migration rates of both these regions.

## g. Median youth unemployment rate per subcontinent

For each subcontinent, calculate the median youth unemployment rate. 
Then create a plot which contains the sub-regions on the y-axis and the median 
unemployment rate on the x-axis.

 - Use a black and white theme (?theme_bw())
 - As geoms use bars. (hint: pay attention to the statistical transformation 
   taking place in geom_bar() – look into argument stat="identity")
 - Color the bars by continent – use a colorblind friendly palette.
 - Make the bars transparent (use alpha = 0.7).
 - Rename the axes.
 - Using fct_reorder from the forcats package, arrange the levels of subcontinent 
   such that in the plot the lowest (bottom) subcontinent contains the lowest median 
   youth unemployment rate and the upper most region contains the highest median 
   youth unemployment rate.
 - Comment on the plot. E.g., what are the regions with the highest vs lowest 
   youth unemployment rate?

**Answer**
```{r}
median_unemployment <- data_tibble %>%
  group_by(Region, Continent) %>%
  summarize(median_youth_unemployment_rate = median(Youth_Unemployment_Rate)) %>%
  ungroup()

median_unemployment <- median_unemployment %>%
  mutate(region = fct_reorder(Region, median_youth_unemployment_rate))

ggplot(median_unemployment, aes(x = median_youth_unemployment_rate, y = region, fill = Continent)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x = "Median Youth Unemployment Rate",
    y = "Subcontinent",
    title = "Median Youth Unemployment Rate by Subcontinent",
    fill = "Continent"
  ) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 10))
```

Youth unemployment is most prevalent in Southern and Northern Africa, but also in Polynesia and Southern Europe. Generally, Asia and Europe score rather low on Youth unemployment, with southern regions tending to perform worse for almost all continents. Interestingly enough, even though most regions of Africa have rather high values, Western Africa belongs to the top performers regarding Youth unemployment.

## h. Median youth unemployment rate per subcontinent – with error bars

The value displayed in the barplot in g. is the result of an aggregation, 
so it might be useful to also plot error bars, to have a general idea on how 
precise the median unemployment is. This can be achieved by plotting the error 
bars which reflect the standard deviation or the interquartile range of the 
variable in each of the subcontinents.

Repeat the plot in h. but include also error bars which reflect the 25% and 75% 
quantiles. You can use geom_errorbar in ggplot2.

**Answer**
```{r}
quantiles_unemployment <- data_tibble %>%
  group_by(Region, Continent) %>%
  summarize(
    median_youth_unemployment_rate = median(Youth_Unemployment_Rate),
    q25 = quantile(Youth_Unemployment_Rate, 0.25),
    q75 = quantile(Youth_Unemployment_Rate, 0.75)
  ) %>%
  ungroup()

quantiles_unemployment <- quantiles_unemployment %>%
  mutate(region = fct_reorder(Region, median_youth_unemployment_rate))

ggplot(quantiles_unemployment, aes(x = median_youth_unemployment_rate, y = region, fill = Continent)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_errorbar(aes(xmin = q25, xmax = q75), width = 0.2) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x = "Median Youth Unemployment Rate",
    y = "Subcontinent",
    title = "Median Youth Unemployment Rate by Subcontinent with Error Bars",
    fill = "Continent"
  ) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 10))
```


## i. Relationship between median age and net migration rate

Using ggplot2, create a plot showing the relationship between 
median age and net migration rate.

 - Color the geoms based on the income status.
 - Add a regression line for each development status (using geom_smooth()).

Comment on the plot. Do you see any relationship between the two variables? Do 
you see any difference among the income levels?

**Answer**
```{r}
ggplot(data_tibble, aes(x = Median_Age, y = Net_Migration_Rate, color = Classification_2020)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Median Age",
    y = "Net Migration Rate",
    title = "Relationship between Median Age and Net Migration Rate",
    color = "Income Status"
  ) +
  theme_minimal() +
  theme(axis.text = element_text(size = 10))

```

The most eye-catching regression can be observed for the Low Income Status group. However, this regression line is clearly heavily influenced by an outlier at the top of the graph. Using a robust regression function could be benefitial here. All other income status classes feature a similar behaviour, with their respective regression lines being almost parallel.

## j. Relationship between youth unemployment and net migration rate

Create a plot as in Task f. but for youth unemployment and net migration rate. 
Comment briefly.

**Answer**
```{r}
ggplot(data_tibble, aes(x = Youth_Unemployment_Rate, y = Net_Migration_Rate, color = Classification_2020)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Youth Unemployment Rate",
    y = "Net Migration Rate",
    title = "Relationship between Youth Unemployment Rate and Net Migration Rate",
    color = "Income Status"
  ) +
  theme_minimal() +
  theme(axis.text = element_text(size = 10))
```

Here, different trends are visible for the low income status group compared with the other ones. For the low income group, a higher youth unemployment rate correlates with a higher net migration rate. The other income status groups tend to the opposite, with higher youth unemployment rates correlating with lower net migration rates.