```{r import_exercise1, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}

import_library <- function(library_name) {
  if (!requireNamespace(library_name, quietly = TRUE)) {
    install.packages(library_name)
  }
  library(library_name, character.only = TRUE)
}

# --------------------------

import_library("ggplot2")
import_library("readr")
import_library("dplyr")
import_library("tidyr")
import_library("tidyverse")
import_library("knitr")
import_library("RColorBrewer")
colorblind_palette <- brewer.pal(4, "Set2")

```

```{r path_determination_exercise1, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

```{r prepare_data_e1}

data_tibble <- read_delim(file = paste0(working_directory_path, "/data/final_data.csv"),
                          show_col_types = FALSE)

data_tibble <- data_tibble %>% 
  filter(Classification_2020 != ".")

data_tibble <- data_tibble %>% 
  mutate(Classification_2020 = factor(Classification_2020, 
                                      levels = c("L", "LM", "UM", "H"), 
                                      labels = c("Low", "Lower-middle", "Upper-middle", "High"), 
                                      ordered = TRUE)
  )

data_tibble <- data_tibble %>%
  mutate(
    Continent = factor(Continent),
    Region = factor(Region)
  )

data_tibble <- data_tibble %>%
 mutate(
    Net_Migration_Rate = na_if(Net_Migration_Rate, ".") %>% as.double(),
    Median_Age = na_if(Median_Age, ".") %>% as.double(),
    Youth_Unemployment_Rate = na_if(Youth_Unemployment_Rate, ".") %>% as.double()
  )

head(data_tibble)

```


## a. Median age in different income levels

Using ggplot2, create a density plot of the median age grouped by income status
groups. The densities for the different groups are superimposed in the same plot
rather than in different plots. Ensure that you order the levels of the income
status such that in the plots the legend is ordered from High (H) to Low (L).

 - The color of the density lines is black.
 - The area under the density curve should be colored differently among 
   the income status levels.
 - For the colors, choose a transparency level of 0.5 for better visibility.
 - Position the legend at the top center of the plot and give it no title 
   (hint: use element_blank()).
 - Rename the x axis as “Median age of population”

Comment briefly on the plot.

**Answer**

```{r a1, dev=c('png')}

# Filter out non-finite values in the Median_Age column
filtered_data <- data_tibble %>%
  filter(is.finite(Median_Age))

# Create the density plot
ggplot(filtered_data, aes(x = Median_Age, fill = Classification_2020)) +
  geom_density(alpha = 0.5, color = "black") +
  scale_fill_manual(values = colorblind_palette) +
  labs(x = "Median age of population") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12))

```

The density plot reveals that low-income countries have predominantly younger
populations, peaking around the 20-25 age range. As income status increases, the
median age of populations also increases, with high-income countries showing the
oldest populations, peaking around 40-45 years. This indicates a trend where
higher-income countries tend to have older populations, while lower-income
countries have younger populations.

## b. Income status in different continents

Investigate how the income status is distributed in the different continents.
 
 - Using ggplot2, create a stacked barplot of absolute frequencies showing how 
   the entities are split into continents and income status. Comment the plot.
 - Create another stacked barplot of relative frequencies 
   (height of the bars should be one). Comment the plot.
 - Create a mosaic plot of continents and income status using base R functions.
 - Briefly comment on the differences between the three plots generated 
   to investigate the income distribution among the different continents.

**Answer**

```{r b1, dev=c('png')}

# Create stacked barplot of absolute frequencies
ggplot(data_tibble, aes(x = Continent, fill = Classification_2020)) +
  geom_bar(position = "stack") +
  scale_fill_manual(values = colorblind_palette) +
  labs(x = "Continent", y = "Count", fill = "Income Status") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12))

```

The absolute frequencies plot shows that Africa predominantly consists of low
and lower-middle-income countries, reflecting its economic challenges, while
Europe and North America are mostly high-income, indicating advanced economic
development. Asia displays a diverse economic landscape with significant
representation across all income categories. Oceania and South America show a
mix of income levels, highlighting varying development within these regions.

```{r b2, dev=c('png')}

# Create stacked barplot of relative frequencies
ggplot(data_tibble, aes(x = Continent, fill = Classification_2020)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = colorblind_palette) +
  labs(x = "Continent", y = "Proportion", fill = "Income Status") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12))

```

The relative frequencies plot makes it easier to compare the proportion of
income statuses within each continent but obscures the actual number of
countries. In contrast, the absolute frequencies plot clearly shows the total
count of countries in each income status category, making it easier to see the
magnitude but harder to compare proportions within continents. Thus, the
relative plot is better for internal distribution comparison, while the absolute
plot is better for understanding total counts.

```{r b3, dev=c('png')}

# Create mosaic plot
mosaicplot(~ Continent + Classification_2020, 
           data = data_tibble, 
           color = colorblind_palette,
           main = "Mosaic Plot of Continents and Income Status",
           xlab = "Continent", 
           ylab = "Income Status")

```

The mosaic plot combines the strengths of both the relative and absolute
frequency plots by showing both the proportions and the total counts of income
statuses within each continent. It uses tile sizes to represent absolute counts
and tile areas to reflect relative proportions, providing a comprehensive view
of the data. This allows for easy comparison of both the distribution balance
within continents and the magnitude of each category.

## c. Income status in different subcontinents

For Asia, investigate further how the income status distribution is in the 
different subcontinents. Use one of the plots in b. for this purpose. 
Comment on the results.

**Answer**

```{r c1, dev=c('png')}

# Filter data to include only Asia
asia_data <- data_tibble %>%
  filter(Continent == "Asia")

# Create stacked barplot of absolute frequencies for Asian subcontinents
ggplot(asia_data, aes(x = Region, fill = Classification_2020)) +
  geom_bar(position = "stack") +
  scale_fill_manual(values = colorblind_palette) +
  labs(x = "Subcontinent", y = "Count", fill = "Income Status") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12))

```

The absolute stacked barplot reveals that Western Asia has the highest diversity
in income statuses, with significant representation in all categories, including
a notable portion in the high-income category. Eastern Asia shows a strong
presence of high-income and upper-middle-income countries, reflecting its
economic development. Central Asia, South-eastern Asia, and Southern Asia are
predominantly lower-middle-income regions, indicating more uniform economic
status within these subcontinents.

## d. Net migration in different continents

 - Using ggplot2, create parallel boxplots showing the distribution of the net 
   migration rate in the different continents.
 - Prettify the plot (change y-, x-axis labels, etc).
 - Identify which country in Asia constitutes the largest negative outlier 
   and which country in Asia constitutes the largest positive outlier.
 - Comment on the plot.

**Answer**

```{r d1, dev=c('png')}

# Filter out non-finite values
filtered_data <- data_tibble %>%
  filter(is.finite(Net_Migration_Rate))

# Create parallel boxplots
ggplot(filtered_data, aes(x = Continent, y = Net_Migration_Rate)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8, outlier.size = 2) +
  labs(x = "Continent", y = "Net Migration Rate") +
  ggtitle("Distribution of Net Migration Rate by Continent") +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5))

# Create parallel boxplots again, 
# but hide the extreme negative outlier
ggplot(filtered_data, aes(x = Continent, y = Net_Migration_Rate)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8, outlier.size = 2) +
  labs(x = "Continent", y = "Net Migration Rate") +
  ggtitle("Distribution of Net Migration Rate by Continent\n(excl. Lebanon)") +
  coord_cartesian(ylim = c(-30, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5))


```

To facilitate the evaluation of the plot, Lebanon was removed in the second
analysis because its extremely low net migration rate skewed the scale, making
the data from the other continents harder to discern.

The boxplot shows net migration rate distributions across continents, excluding
extreme negative outliers. Asia and Africa have the most variation, with many
outliers, indicating diverse migration dynamics. This suggests different
countries experience varying economic, political, and social conditions. Europe,
North America, and Oceania have more centralized distributions with fewer
outliers, reflecting stable migration trends due to established economic
conditions and policies. South America shows a narrower range and fewer extreme
values, indicating more uniform migration trends within the continent.


```{r d2, dev=c('png')}

# Filter data for Asia and remove non-finite values
asia_data <- data_tibble %>%
  filter(Continent == "Asia" & is.finite(Net_Migration_Rate))

# Identify the largest negative and positive outliers in Asia
largest_negative_outlier <- asia_data %>%
  filter(Net_Migration_Rate == min(Net_Migration_Rate, na.rm = TRUE)) %>%
  select(Country, Net_Migration_Rate)

largest_positive_outlier <- asia_data %>%
  filter(Net_Migration_Rate == max(Net_Migration_Rate, na.rm = TRUE)) %>%
  select(Country, Net_Migration_Rate)

# Display the results
print(largest_negative_outlier)
print(largest_positive_outlier)

```

## e. Net migration in different subcontinents

The graph in d. clearly does not convey the whole picture. 
It would be interesting also to look at the subcontinents, as it is likely that 
a lot of migration flows happen within the continent.

 - Investigate the net migration in different subcontinents 
   using again parallel boxplots. 
   Group the boxplots by continent (hint: use facet_grid with scales = "free_x").
 - Remember to prettify the plot (rotate axis labels if needed).
 - Describe what you see.

**Answer**

```{r e1, dev=c('png')}

# Filter out non-finite values
filtered_data <- data_tibble %>%
  filter(is.finite(Net_Migration_Rate))

# Create parallel boxplots grouped by continent
ggplot(filtered_data, aes(x = Region, y = Net_Migration_Rate)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8, outlier.size = 2) +
  labs(x = "Subcontinent", y = "Net Migration Rate") +
  ggtitle("Distribution of Net Migration Rate by Subcontinent") +
  coord_cartesian(ylim = c(-30, 30)) +
  facet_grid(. ~ Continent, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        strip.text.x = element_text(size = 12, face = "bold"))

```

When splitting continents into subcontinents, the detailed breakdown reveals
significant regional variations within each continent that are not visible in
the broader view. Subcontinent-specific migration patterns and outliers become
apparent, such as the distinct differences between Central and Western Asia or
among the subregions of Africa.
