```{r import_exercise2, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}

import_library <- function(library_name) {
  if (!requireNamespace(library_name, quietly = TRUE)) {
    install.packages(library_name)
  }
  library(library_name, character.only = TRUE)
}

# --------------------------

import_library("ggplot2")
import_library("readr")
import_library("dplyr")
import_library("tidyr")
import_library("tidyverse")
import_library("knitr")
import_library("forcats")

```

```{r path_determination_exercise2, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

## f. Median net migration rate per subcontinent.

The plot in task e. shows the distribution of the net migration rate for each subcontinent. 
Here you will work on visualizing only one summary statistic, namely the median.
For each subcontinent, calculate the median net migration rate. Then create a plot which contains the
sub-regions on the y-axis and the median net migration rate on the x-axis.

 - As geoms use points.
 - Color the points by continent – use a colorblind friendly palette 
   (see e.g., [here](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)).
 - Rename the axes.
 - Using fct_reorder from the forcats package, arrange the levels of subcontinent 
   such that in the plot the lowest (bottom) subcontinent contains the lowest 
   median net migration rate and the upper most region contains the highest median 
   net migration rate.
 - Comment on the plot. E.g., what are the regions with the most influx? 
   What are the regions with the most outflux?

**Answer**

```{r f1, dev=c('png')}

```

## g. Median youth unemployment rate per subcontinent

For each subcontinent, calculate the median youth unemployment rate. 
Then create a plot which contains the sub-regions on the y-axis and the median 
unemployment rate on the x-axis.

 - Use a black and white theme (?theme_bw())
 - As geoms use bars. (hint: pay attention to the statistical transformation 
   taking place in geom_bar() – look into argument stat="identity")
 - Color the bars by continent – use a colorblind friendly palette.
 - Make the bars transparent (use alpha = 0.7).
 - Rename the axes.
 - Using fct_reorder from the forcats package, arrange the levels of subcontinent 
   such that in the plot the lowest (bottom) subcontinent contains the lowest median 
   youth unemployment rate and the upper most region contains the highest median 
   youth unemployment rate.
 - Comment on the plot. E.g., what are the regions with the highest vs lowest 
   youth unemployment rate?

**Answer**

## h. Median youth unemployment rate per subcontinent – with error bars

The value displayed in the barplot in g. is the result of an aggregation, 
so it might be useful to also plot error bars, to have a general idea on how 
precise the median unemployment is. This can be achieved by plotting the error 
bars which reflect the standard deviation or the interquartile range of the 
variable in each of the subcontinents.

Repeat the plot in h. but include also error bars which reflect the 25% and 75% 
quantiles. You can use geom_errorbar in ggplot2.

**Answer**

## i. Relationship between median age and net migration rate

Using ggplot2, create a plot showing the relationship between 
median age and net migration rate.

 - Color the geoms based on the income status.
 - Add a regression line for each development status (using geom_smooth()).

Comment on the plot. Do you see any relationship between the two variables? D
o you see any difference among the income levels?

**Answer**

## j. Relationship between youth unemployment and net migration rate

Create a plot as in Task f. but for youth unemployment and net migration rate. 
Comment briefly.

**Answer**
