## 2. Data analysis - Part 1

```{r import_exercise2, message=FALSE, warning=FALSE, include=FALSE}

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

if (!requireNamespace("tidyr", quietly = TRUE)) {
  install.packages("tidyr")
}
library(tidyr)

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
library(tidyverse)

if (!requireNamespace("knitr", quietly = TRUE)) {
  install.packages("knitr")
}
library(knitr)

```

```{r path_determination_exercise2, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

```{r init, echo=FALSE}
# load the data
df_vars <- read.csv("../data/df_vars.csv")
```
### g.

Make a frequency table for the status variable in the merged data set. Briefly
comment on the results.

```{r g}
df_vars %>% mutate(status = fct_infreq(Classification_2020)) %>% count(status)
```
We can see that the biggest part of the countries actually falls into the high income class, in the middle classes "UM" and "LM" there are about an even amount of countries. Only 27 countries fall under low income. However there are relatively many NA values present which might skew the previous observations.

### h.

What is the distribution of income status in the different continents? Compute
the absolute frequencies as well as the relative frequency of status within each
continent. Briefly comment on the results.

```{r h}
by_cont <- df_vars %>% group_by(Continent)
income_status <- by_cont %>% count(Classification_2020) %>% mutate(abs = n, freq = n/sum(n))
income_status$n <- NULL
income_status
```
From this table it is clearly visible that the continents with the highes income are Europe and North America, both of which have no countries classified in "L" and very few in "LM". Next is Oceania also having no "L"-countries, but an even spread between the other categories. South America is also void of "L"-classified countries, with most of them being in the "UM" category. Asia has a few "L"-countries and also an even spread between the others. The poorest continent is Africa, only having one country in the "H" category and many in the "L" and "LM" categories.

### i.

From h. identify the countries which are the only ones in their respective
group. Explain in few words the output.

```{r i}
abs_1 <- income_status %>% filter(abs==1) %>% select(Continent, Classification_2020)
abs_1 <- left_join(abs_1, df_vars)
abs_1 %>% select(Continent, Classification_2020, Country)
```
In the output we have 1 NA value, Antarctica (which is the only country on the continent), Seychelles (the only "H" rated country in Africa) and Ukraine and Bolivia which are the only "LM" rated countries in Europe/South America respectively.

### j.

For each continent count the number of sub-regions in the data set. How granular
are the subcontinents that you employ in the analysis?

```{r j}
by_cont %>% summarise(sub_regions = n_distinct(Region))
```

The continents are separated into sub regions in a very different way. None of them have more than 5 sub-regions, however while a rather small continent such as Europe has 4 sub-regions, really big ones like Asia and Africa are also only split into 5. South America even has only 1 sub-region.

### k.

Look at the frequency distribution of income status in the subregions of North- 
and South-Americas. Comment on the results.

```{r k}
na <- df_vars %>% filter(Continent=="North America")
sa <- df_vars %>% filter(Continent=="South America")
americas <- full_join(na, sa)

by_region <- americas %>% group_by(Region)
income_status_am <- by_region %>% count(Classification_2020) %>% mutate(abs = n, freq = n/sum(n))
income_status_am$n <- NULL
income_status_am
```

The Caribbean seems to have rather high income with the biggest part of countries being classified in "H", the same is true for North America. Central America is equally split between "LM" and "UM". South American countries are mostly classified in "UM", with a few in "H" and "LM".

### l.

Dig deeper into the low-middle income countries of the Americas. 
Which ones are they? Are they primarily small island states in the Caribbean? 
Comment.

```{r l}
americas_LM <- income_status_am %>% filter(Classification_2020=="LM")
americas_LM <- left_join(americas_LM, df_vars)
americas_LM
```

They are primarily states in Central America, only one of them is a Central American island nation.