## 2. Data analysis - Part 1

```{r import_exercise2, message=FALSE, warning=FALSE, include=FALSE}

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

if (!requireNamespace("tidyr", quietly = TRUE)) {
  install.packages("tidyr")
}
library(tidyr)

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
library(tidyverse)

if (!requireNamespace("knitr", quietly = TRUE)) {
  install.packages("knitr")
}
library(knitr)

```

```{r path_determination_exercise2, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

### g.

Make a frequency table for the status variable in the merged data set. Briefly
comment on the results.

**Answer**

### h.

What is the distribution of income status in the different continents? Compute
the absolute frequencies as well as the relative frequency of status within each
continent. Briefly comment on the results.

**Answer**

### i.

From h. identify the countries which are the only ones in their respective
group. Explain in few words the output.

**Answer**

### j.

For each continent count the number of sub-regions in the data set. How granular
are the subcontinents that you employ in the analysis?

**Answer**

### k.

Look at the frequency distribution of income status in the subregions of Nort- 
and South-Americas. Comment on the results.

**Answer**

### l.

Dig deeper into the low-middle income countries of the Americas. 
Which ones are they? Are they primarily small island states in the Caribbean? 
Comment.

**Answer**