## 3. Data analysis - Part 2

```{r import_exercise3, message=FALSE, warning=FALSE, include=FALSE}

if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

if (!requireNamespace("tidyr", quietly = TRUE)) {
  install.packages("tidyr")
}
library(tidyr)

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
library(tidyverse)

if (!requireNamespace("knitr", quietly = TRUE)) {
  install.packages("knitr")
}
library(knitr)

```

```{r path_determination_exercise3, include=FALSE}

# Determine if it's running as a child or not
# to fix paths when knitting
if (requireNamespace("rstudioapi", quietly = TRUE)) {
 script_path <- rstudioapi::getActiveDocumentContext()$path
 
 if (endsWith(dirname(script_path), "exercises")) {
   working_directory_path <- dirname(dirname(script_path))
 } else {
   working_directory_path <- dirname(script_path)
 }
 
} else {
 message("RStudio API is not available")
}

```

### m.

Create a table of average values for median age, youth unemployment rate and net
migration rate separated into income status. Make sure that in the output, the
ordering of the income classes is proper (i.e., L, LM, UM, H or the other way
around). Briefly comment the results.

**Answer**

### n.

Look also at the standard deviation instead of the mean in m. 
Do you gain additional insights? Briefly comment the results.

**Answer**

### o.

Repeat the analysis in $\mathrm{m}$. for each income status and continent combination. 
Discuss the results.

**Answer**

### p.

Identify countries which are doing well in terms of both youth unemployment and
net migration rate (in the top $25 \%$ of their respective continent in terms of
net migration rate and in the bottom $25 \%$ of their respective continent in
terms of youth unemployment).

**Answer**

### r.

Export the final data set to a csv with ";" separator and "." as a symbol for
missing values; no rownames should be included in the csv. Upload the .csv to
TUWEL together with your .Rmd and .html (or .pdf).

**Answer**

